Floral resource choice by bumble bee species in the Berchtesgadener Alps

### 2020-12-07 analyses/old_scripts/niche_relations_v6.Rmd
update lab desktop

  scores <- y %>% 
    #filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")
    filter((abs(MDS1) > 0.1 | abs(MDS2) > 0.1) | Score == "BB.sp")

@@

  x %>%
    filter(caste != "male") %>%
    #filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
    filter(!bb.sp %in% c("gers", "humi")) %>% # gers because it's a specialist, humi because it's so rare
    #filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    group_by(site, date, bb.sp, !!criterion, ...) %>% # pooling all sites and years
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(site, date, bb.sp) %>%
    mutate(prop = freq/sum(freq)) %>%
    group_by(site, date, ...) %>%
    complete(bb.sp, !!criterion, fill = list(prop = 0)) %>% # fill in zeros for unvisited plants
    group_by(..., bb.sp, !!criterion) %>%
    summarize(prop = median(prop)) %>%
    #dplyr::select(-freq) %>% # drop freq variable
    group_by(...) %>%
    nest()  %>% # nest into list-column
    mutate(comm1 = map(data, # spread into the kind of array vegan likes
                       function(x) pivot_wider(x, names_from = !!criterion, values_from = prop) %>%
                         replace(is.na(.), 0)
                       ), 
           comm2 = map(comm1, # column to rowname
                       function(x) column_to_rownames(x, var = "bb.sp")
                       ),
           pcoa = map(comm2, # call PCoA
                      function(x) capscale(x ~ 1, distance = "horn")
                      ),
           traits = map(comm1, # join BB traits
                        function(x) x %>% 
                          left_join(y) %>% # y is the trait table, an argument in the main function call
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.w) # pbl.w is the variable I'm interested in
                        ),
           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
                         function(x, y) adonis(x ~ pbl.w, data = y, method = "horn")
                         ),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           adonis.r2 = map(adonis, # extract adonis r2
                           function(x) x$aov.tab$R2[[1]]
                           ),
           adonis.p = map(adonis, # extract adonis p-value
                          function(x) x$aov.tab$'Pr(>F)'[[1]]
                          ),
           ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 8)),
           #ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 5)),
           ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
           ordisurf.p = map(ordisurf, function(x) summary(x)$s.pv),
           ordisurf.plot = map2(ordisurf, pcoa.fort, surfplot),
           var.exp.MDS1 = map(pcoa, # extract variance explained by MDS1
                      function(x) x$CA$eig[[1]] / sum(x$CA$eig)
                      ), 
           var.exp.MDS2 = map(pcoa, # extract variance explained by MDS2
                      function(x) x$CA$eig[[2]] / sum(x$CA$eig)
                      )) %>%
    unnest(cols = c(ordisurf.dev_exp, ordisurf.p, adonis.r2, 
                    adonis.p, var.exp.MDS1, var.exp.MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
    mutate(var.exp.MDS1_MDS2 = var.exp.MDS1 + var.exp.MDS2) %>%
    select(..., pcoa, pcoa.plot, var.exp.MDS1_MDS2, var.exp.MDS1, var.exp.MDS2, 
            ordisurf.dev_exp, ordisurf.p, adonis.r2, adonis.p,
            ordisurf, ordisurf.plot, adonis, comm2, traits, pcoa.fort)
}

@@

# Morphological
niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.ss, elev.class2)

@@

## 1.6 Tree plots
This visualization deserves to be pulled out as a stand-alone chunk. It's so good.
```{r}
### Visitation by BB species, parsed by elev.class2 and period
visitation_abund.sp <- net %>%
  filter(caste != "male") %>%
  filter(!bb.sp %in% c("gers", "humi")) %>% # only focal species
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
  # mutate(k.type.ss = factor(k.type.ss),
  #        color = factor(color),
  #        bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         color = factor(color),
         bb.sp = factor(bb.sp)) %>%
  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class2, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>% 
  left_join(bb_traits) %>%
  arrange(elev.class2, period, bb.sp, -prop)

>>>>

  scores <- y %>% 
    filter((abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")

@@

  x %>%
    filter(caste != "male") %>%
    #filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
    filter(!bb.sp %in% c("gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
    group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(..., bb.sp) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    group_by(...) %>%
    nest() %>% # nest into list-column
    mutate(comm1 = map(data, # spread into the kind of array vegan likes
                       function(x) pivot_wider(x, names_from = !!criterion, values_from = prop) %>%
                         replace(is.na(.), 0)
                       ), 
           comm2 = map(comm1, # column to rowname
                       function(x) column_to_rownames(x, var = "bb.sp")
                       ),
           pcoa = map(comm2, # call PCoA
                      function(x) capscale(x ~ 1, distance = "horn")
                      ),
           traits = map(comm1, # join BB traits
                        function(x) x %>% 
                          left_join(y) %>% # y is the trait table, an argument in the main function call
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.w) # pbl.w is the variable I'm interested in
                        ),
           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
                         function(x, y) adonis(x ~ pbl.w, data = y, method = "horn")
                         ),
           pcoa.fort = map(pcoa, pcoa_fort),
           pcoa.plot = map(pcoa.fort, pcoa_plot),
           adonis.r2 = map(adonis, # extract adonis r2
                           function(x) x$aov.tab$R2[[1]]
                           ),
           adonis.p = map(adonis, # extract adonis p-value
                          function(x) x$aov.tab$'Pr(>F)'[[1]]
                          ),
           ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 8)),
           ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
           ordisurf.p = map(ordisurf, function(x) summary(x)$s.pv),
           ordisurf.plot = map2(ordisurf, pcoa.fort, surfplot),
           var.exp.MDS1 = map(pcoa, # extract variance explained by MDS1
                      function(x) x$CA$eig[[1]] / sum(x$CA$eig)
                      ), 
           var.exp.MDS2 = map(pcoa, # extract variance explained by MDS2
                      function(x) x$CA$eig[[2]] / sum(x$CA$eig)
                      )) %>%
    unnest(cols = c(ordisurf.dev_exp, ordisurf.p, adonis.r2, 
                    adonis.p, var.exp.MDS1, var.exp.MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
    mutate(var.exp.MDS1_MDS2 = var.exp.MDS1 + var.exp.MDS2) %>%
    select(..., pcoa, pcoa.plot, var.exp.MDS1_MDS2, var.exp.MDS1, var.exp.MDS2, 
            ordisurf.dev_exp, ordisurf.p, adonis.r2, adonis.p,
            ordisurf, ordisurf.plot, adonis, comm2, traits, pcoa.fort)
}

@@

# Morphological
niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2)

@@

## 1.6 Tree plots
This visualization deserves to be pulled out as a stand-alone chunk. It's so good.
```{r}
### Visitation by BB species, parsed by elev.class2 and period
visitation_abund.sp <- net %>%
  filter(caste != "male") %>%
  filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # only focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 3, labels = c("early",  "mid", "late"))) %>%
  # mutate(k.type.ss = factor(k.type.ss),
  #        color = factor(color),
  #        bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         color = factor(color),
         bb.sp = factor(bb.sp)) %>%
  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class2, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>% 
  left_join(bb_traits) %>%
  arrange(elev.class2, period, bb.sp, -prop)
